I am grateful to my family for their encouragement during my entire education.
There have been a lot of ups and downs and they have always supported me in my decisions.

Thank you to my advisor, Kyunghyun Cho, who has been there for me when I needed him.
He pushed me to think more clearly and critically about my ideas, but more importantly, he is one of the best people I know.
Thanks are also due to Martin Riedmiller and Abhinav Gupta for mentoring me on important projects and opening their groups to me, as well as for being on my committee.
I am also grateful to Yann Lecun and Lerrel Pinto for serving on my committee and for their valuable insights.
Tejas Kulkarni and Josh Tenenbaum will always deserve thanks for taking me in during my Master's and introducing me to the world of machine learning research.

I have been lucky to collaborate with wonderful, gifted people during my PhD, including David Brandfonbrener, Michael Bloesch, Jost Tobias Springenberg, Abbas Abdolmaleki, Mikael Henaff, Min Jae Song, Rajesh Ranganath, Joan Bruna, Jaan Altosaar, Owen Lewis, Emily Denton, Rob Fergus, and Rajat Agarwal, as well as those mentioned above.
Without all of you this would never have gotten done, and I would not have become the researcher I am.

My house full of machine learning people has been a continual pleasure.
I would like to thank everyone who lived there, especially Cinjon, David, Elman, and Martin, for being around to spitball ideas or even, every once in a while, to take my mind off of research.

Finally, thank you to Margaret for making the last year of my PhD a delightful adventure instead of a stressful slog.
It's possible that I could do it without you, but why would I want to?
